The Cost of Exclusion: Political Discrimination and Trust Deficits in the European Union

### 2025-07-10 article/main.tex
Updated article

While this framework has typically focused on ethnic or cultural exclusion (Oskoii, 2016; Wilkes & Wu, 2019; Tyrberg, 2023), political exclusion—based on one's beliefs or partisan identity—has received comparatively less attention. Yet, discrimination or harassment for political reasons may function in similar ways. Such experiences act as signals of exclusion from public discourse or decision-making, leading individuals to feel marginalized from the political community. This sense of exclusion can cause individuals to question the impartiality, responsiveness, and integrity of institutions.

@@

Political dissonance is exacerbated in polarized societies where outgroup hostility, moral condemnation, and dehumanization are common (Iyengar & Westwood, 2015). This intensifies the emotional toll of political discrimination and heightens perceptions of institutional betrayal (Hetherington \& Rudolph, 2015). In turn, this reduces trust and discourages political participation (Michelson, 2003; Lelkes \& Westwood, 2017).

@@

>>>>

While this framework has typically focused on ethnic or cultural exclusion (Oskoii, 2016; Wilkes \& Wu, 2019; Tyrberg, 2023), political exclusion—based on one's beliefs or partisan identity—has received comparatively less attention. Yet, discrimination or harassment for political reasons may function in similar ways. Such experiences act as signals of exclusion from public discourse or decision-making, leading individuals to feel marginalized from the political community. This sense of exclusion can cause individuals to question the impartiality, responsiveness, and integrity of institutions.

@@

Political dissonance is exacerbated in polarized societies where outgroup hostility, moral condemnation, and dehumanization are common (Iyengar \& Westwood, 2015). This intensifies the emotional toll of political discrimination and heightens perceptions of institutional betrayal (Hetherington \& Rudolph, 2015). In turn, this reduces trust and discourages political participation (Michelson, 2003; Lelkes \& Westwood, 2017).

@@

% \newpage
% \printbibliography
